API design for machine learning software: experiences from the scikit-learn project

### 2013-06-26 paper.tex
inputs -> input, outputs -> output
I prefer to use these as uncountable nouns; did that where it looked
better than the countable usage.

Finally, predictors should also provide a \texttt{score} function to assess the
performance on some input data. In supervised learning estimators, this method
takes as inputs arrays \texttt{X\_test} and \texttt{y\_test} and typically
computes the coefficient of determination between \texttt{y\_test} and
\texttt{predict(X\_test)} in regression, or the accuracy
in classification. From an API point of view,
the only guideline is that the \texttt{score} method returns a value
that quantifies the quality of its predictions (the higher, the better).
Unsupervised learning estimators may also exposes a \texttt{score} function
to compute, for instance, the likelihood of the given data under its model.

@@

Since it is common to modify or filter data before feeding it to a learning
algorithm, some estimators in the library implement a \textit{transformer}
interface which defines a \texttt{transform} method. It takes as input some new
data \texttt{X\_test} and yields as output a transformed version of
\texttt{X\_test}. Preprocessing, feature selection and dimensionality reduction
algorithms are all provided as transformers within the library.  In our example,
if one wanted to preprocess the inputs \texttt{X\_train} before fitting the
logistic regression estimator, e.g., by standardizing the feature values, one
could thus simply write:
\begin{pythoncode}
from sklearn.preprocessing import StandardScaler

@@

\texttt{FeatureUnion} objects combine multiple transformers into a single one
that concatenates their outputs. A union of two transformers that
map inputs having $d$ features to $d'$ and $d''$ features respectively is
a transformer that maps its $d$ input features to $d' + d''$ features.
This generalizes in the obvious way to more than two transformers.
In terms of API, a \texttt{FeatureUnion} takes as input a list of transformers.
Calling \texttt{fit} on the union is the same as calling \texttt{fit}
independently on each of the transformers and then joining their outputs.

@@

In \sklearn, model selection is supported in two distinct meta-estimators,
\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as inputs an
estimator (basic or composite), whose hyper-parameters must be optimized, and a
hyper-parameter grid, which defines the values to search. Optionally, they can
also take a cross-validation scheme and a score function.  The library provides
various such cross-validation schemes, including $k$-fold (default),
leave-one-out and iterated shuffled $k$-folds with optional sub-sampling.
The score function used by default is the \texttt{score} method
provided by the estimator. However, the library provides a large variety of
score functions that the user can choose from,
including accuracy, AUC, $F_1$ score (classification),
$R^2$ score and mean squared error (regression).

@@

In other places of the library, in particular in the vectorization code
for unstructured inputs, the toolkit is also designed to be
extensible. Here, estimators provide hooks for user-defined code: objects or
functions that provide a specific API can be given as arguments at vectorizer
construction time. The library then calls into this code, communicating with it by passing objects of standard Python/Numpy types.
Again, such external user code can be kept agnostic of the \sklearn
class hierarchy.

@@

Recent years have witnessed a rising interest in machine learning and data mining
with applications in many fields.
With this rise comes a host of machine learning packages
(both open source and proprietary) with which \sklearn competes.
Although it can also be used through its Java API, Weka \citep{hall2009weka} is
popularly used through its graphical user interface (GUI). We argue that a GUI
lacks flexibility, makes experiments difficult to reproduce and has a high
maintenance cost from a development point of view. For these reasons, we make no
attempt to provide such a GUI in \sklearn.
Other existing machine learning packages 
such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
and Vowpal~Wabbit\footnote{\url{http://hunch.net/~vw/}}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process inputs/outputs,
and will do so in a variety of languages.
We argue that this makes it harder for researchers and developers
to exchange and collaborate on software.

>>>>

Finally, predictors should also provide a \texttt{score} function to assess the
performance on some input data. In supervised learning estimators, this method
takes as input arrays \texttt{X\_test} and \texttt{y\_test} and typically
computes the coefficient of determination between \texttt{y\_test} and
\texttt{predict(X\_test)} in regression, or the accuracy
in classification. From an API point of view,
the only guideline is that the \texttt{score} method returns a value
that quantifies the quality of its predictions (the higher, the better).
Unsupervised learning estimators may also exposes a \texttt{score} function
to compute, for instance, the likelihood of the given data under its model.

@@

Since it is common to modify or filter data before feeding it to a learning
algorithm, some estimators in the library implement a \textit{transformer}
interface which defines a \texttt{transform} method. It takes as input some new
data \texttt{X\_test} and yields as output a transformed version of
\texttt{X\_test}. Preprocessing, feature selection and dimensionality reduction
algorithms are all provided as transformers within the library.  In our example,
if one wanted to preprocess the input \texttt{X\_train} before fitting the
logistic regression estimator, e.g., by standardizing the feature values, one
could thus simply write:
\begin{pythoncode}
from sklearn.preprocessing import StandardScaler

@@

\texttt{FeatureUnion} objects combine multiple transformers into a single one
that concatenates their outputs. A union of two transformers that
map input having $d$ features to $d'$ and $d''$ features respectively is
a transformer that maps its $d$ input features to $d' + d''$ features.
This generalizes in the obvious way to more than two transformers.
In terms of API, a \texttt{FeatureUnion} takes as input a list of transformers.
Calling \texttt{fit} on the union is the same as calling \texttt{fit}
independently on each of the transformers and then joining their outputs.

@@

In \sklearn, model selection is supported in two distinct meta-estimators,
\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as input an
estimator (basic or composite), whose hyper-parameters must be optimized, and a
hyper-parameter grid, which defines the values to search. Optionally, they can
also take a cross-validation scheme and a score function.  The library provides
various such cross-validation schemes, including $k$-fold (default),
leave-one-out and iterated shuffled $k$-folds with optional sub-sampling.
The score function used by default is the \texttt{score} method
provided by the estimator. However, the library provides a large variety of
score functions that the user can choose from,
including accuracy, AUC, $F_1$ score (classification),
$R^2$ score and mean squared error (regression).

@@

In other places of the library, in particular in the vectorization code
for unstructured input, the toolkit is also designed to be
extensible. Here, estimators provide hooks for user-defined code: objects or
functions that provide a specific API can be given as arguments at vectorizer
construction time. The library then calls into this code, communicating with it by passing objects of standard Python/Numpy types.
Again, such external user code can be kept agnostic of the \sklearn
class hierarchy.

@@

Recent years have witnessed a rising interest in machine learning and data mining
with applications in many fields.
With this rise comes a host of machine learning packages
(both open source and proprietary) with which \sklearn competes.
Although it can also be used through its Java API, Weka \citep{hall2009weka} is
popularly used through its graphical user interface (GUI). We argue that a GUI
lacks flexibility, makes experiments difficult to reproduce and has a high
maintenance cost from a development point of view. For these reasons, we make no
attempt to provide such a GUI in \sklearn.
Other existing machine learning packages 
such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
and Vowpal~Wabbit\footnote{\url{http://hunch.net/~vw/}}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process input/output,
and will do so in a variety of languages.
We argue that this makes it harder for researchers and developers
to exchange and collaborate on software.
